### 2015-03-16 staticanalysis.tex
updates in the static analysis section

For each entry point $\mathit{ep}_{i}$, we compute its lockset $\mathit{LS}_{i}$ that associates each memory location $m$ to the set of locks that are always held when $m$ is accessed during the execution of $\mathit{ep}_{i}$. Furthermore, for each entry point $\mathit{ep}_{i}$, we compute its read-set $R_{i}\lbrack m\rbrack \rightarrow \{true, false\}$ and its write-set $W_{i}\lbrack m\rbrack \rightarrow \{true, false\}$, which map each $m$ to true if and only if $\mathit{ep}_{i}$ reads or writes respectively to $m$ during some execution. In the actual implementation we avoid the use of quantifiers by using a different lockset, read-set and write-set for each shared memory location. \PDComment{any concrete reference that quantifiers might be harder for SMT solving?}

@@

The main limitation of \whoop is that it is based on static lockset analysis. As an example of this limitation is that we do not check for dynamically created locks (or for locks that are provided by external libraries). Because we perform two thread reduction, we also over-approximate the read accesses, which is another source of false positives. Environmental modelling is also another limitation of \whoop. As aforementioned, enriching \whoop with domain-specific knowledge to increase precision is an ongoing manual effort, but requires expertise in the Linux kernel.

We argue that further increasing the precision of our models is orthogonal to the contributions of this paper. Moreover, despite unavoidable imprecisions during the sound static data race analysis, \whoop can use the reported information to significantly speedup precise bug-finders for concurrent programs, as seen in Sections~\ref{bugfinding} and~\ref{evaluation}.

>>>>

For each entry point $\mathit{ep}_{i}$, we compute its lockset $\mathit{LS}_{i}$ that associates each memory location $m$ to the set of locks that are always held when $m$ is accessed during the execution of $\mathit{ep}_{i}$. Furthermore, for each entry point $\mathit{ep}_{i}$, we compute its read-set $R_{i}\lbrack m\rbrack \rightarrow \{true, false\}$ and its write-set $W_{i}\lbrack m\rbrack \rightarrow \{true, false\}$, which map each $m$ to true if and only if $\mathit{ep}_{i}$ reads or writes respectively to $m$ during some execution. In the actual implementation we avoid the use of quantifiers by using a different lockset, read-set and write-set for each shared memory location.

@@

\whoop inherits all limitations of lockset analysis (as discussed in Section~\ref{bg:lockset}): a violation of the locking discipline does not always correspond to a real error (e.g. \whoop does not reasons about lockfree data structures in a special way, it just over-approximates their read accesses); and we do not check for dynamically created locks (or for locks that are provided by external libraries). Because we perform two thread reduction, we also over-approximate the read accesses, which is another source of false positives.

Environmental modelling is another limitation of \whoop. As aforementioned, enriching \whoop with domain-specific knowledge to increase precision is an ongoing manual effort, but requires expertise in the Linux kernel. We argue that further increasing the precision of our models is orthogonal to the contributions of this paper. Moreover, despite unavoidable imprecisions during the sound static data race analysis, \whoop can use the reported information to significantly speedup precise bug-finders for concurrent programs, as seen in Sections~\ref{bugfinding} and~\ref{evaluation}.
